---
title: "Process"
---

<div class="alert alert-danger" role="alert">
  <strong> NOTE: </strong> Testing data for K-Means Clustering 🧪🧪🧪
</div>

<details open="open">
  <summary>Table of Contents</summary>
  <ol>
    <li><a href="#dependency-imports">Dependencies</a></li>
    <li><a href="#importing-pisa-2018-data">Importing Data</a></li>
    <li><a href="#exploring-the-data">Data Exploration</a></li>
    <li><a href="#data-cleaning">Data Cleaning</a></li>
    <li><a href="#data-rocessing">Data Preprocessing</a></li>
    <li><a href="#acknowledgements">Acknowledgements</a></li>
  </ol>
</details>

<hr/>

### Dependency Imports
```{r message=FALSE}
library(learningtower)
library(dplyr)
library(tidyr)
library(factoextra)
library(reactable)
```

<hr/>

### Importing PISA 2018 Data
```{r message=FALSE, cache.lazy=TRUE}

# Loading the PISA 2018 data and assigning it to variable PISA_2018
PISA_2018 <- load_student(2018)

# Sampling 10 observations from the data-set for showing example
PISA_2018_SAMPLE <- sample_n(PISA_2018, 10)

# Output of the sample
reactable(PISA_2018_SAMPLE, 
  searchable = TRUE,
  striped = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  theme = reactableTheme(
    borderColor = "#dfe2e5",
    stripedColor = "#f6f8fa",
    highlightColor = "#f0f5f9",
    cellPadding = "8px 12px"))
```

#### PISA 2018 Data Format

<div class="alert alert-info" role="alert">
  <strong> Total Number of Observations: </strong> 612,004 in the year 2018 alone
</div>
<table class="table table-hover">
  <thead>
    <tr>
      <th>Variable Name</th>
      <th>Variable Type</th>
    </tr>
  </thead>
  <tbody>
    <tr>
      <td>year</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>country</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>school_id</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>student_id</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>mother_educ</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>father_educ</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>gender</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>computer</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>internet</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>math</td>
      <td>Numeric</td>
    </tr>
    <tr>
      <td>science</td>
      <td>Numeric</td>
    </tr>
    <tr>
      <td>read</td>
      <td>Numeric</td>
    </tr>
    <tr>
      <td>stu_wgt</td>
      <td>Numeric</td>
    </tr>
    <tr>
      <td>desk</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>room</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>dishwasher</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>television</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>computer_n</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>car</td>
      <td>Factor</td>
    </tr>
  <tbody>
</table>

<hr/>

### Exploring the Data 

```{r message=FALSE}
# Class of the data
class(PISA_2018)

# Total number of rows and columns in the data-set
dim(PISA_2018)
```

<table class="table table-hover">
  <tr>
    <th>
      Class
    </th>
    <th>
      Dimension
    </th>
  </tr>
  <tr>
    <td align="center">
      data.frame
    </td>
    <td align="center">
      612004 Observations
      <br/>
      22 Variables
    </td>
  </tr>
</table>

#### Summary Statistics of the Data-set
```{r message=FALSE}
# Mathematics summary Stats
MATH_STAT_2018 <- summary(PISA_2018$math)
```

<table class="table table-hover">
  <tr>
    <th>
      Mininum
    </th>
    <th>
      1st Quarter
    </th>
    <th>
      Median
    </th>
    <th>
      Mean
    </th>
    <th>
      3rd Quarter
    </th>
    <th>
      Max
    </th>
    <th>
      Missing
    </th>
  </tr>
  <tr>
    <td align="center">
      24.74
    </td>
    <td align="center">
      386.66
    </td>
    <td align="center">
      460.66
    </td>
    <td align="center">
      461.22
    </td>
    <td align="center">
      535.28
    </td>
    <td align="center">
      888.06
    </td>
    <td align="center">
      5377
    </td>
  </tr>
</table>

```{r message=FALSE}
# Science summary Stats
SCI_STAT_2018 <- summary(PISA_2018$science)
```

<table class="table table-hover">
  <tr>
    <th>
      Mininum
    </th>
    <th>
      1st Quarter
    </th>
    <th>
      Median
    </th>
    <th>
      Mean
    </th>
    <th>
      3rd Quarter
    </th>
    <th>
      Max
    </th>
    <th>
      Missing
    </th>
  </tr>
  <tr>
    <td align="center">
      58.74
    </td>
    <td align="center">
      385.60
    </td>
    <td align="center">
      458.20
    </td>
    <td align="center">
      460.69
    </td>
    <td align="center">
      533.44
    </td>
    <td align="center">
      886.08
    </td>
    <td align="center">
      5377
    </td>
  </tr>
</table>
```{r message=FALSE}
  PISA_MATH_2018 <- PISA_2018$math
  hist(PISA_MATH_2018, 
       main = paste("Histogram of", "PISA 2018 Mathematics Performance"), 
       col="#3e3f3a")
```

<div class="alert alert-info" role="alert">
  <strong> Distribution: </strong> Normal
</div>

```{r message=FALSE}
  PISA_SCI_2018 <- PISA_2018$science
  hist(PISA_SCI_2018, 
       main = paste("Histogram of", "PISA 2018 Science Performance"),
       col = "#3e3f3a")
```

<div class="alert alert-info" role="alert">
  <strong> Distribution: </strong> Normal
</div>

```{r message=FALSE}
  boxplot(na.omit(PISA_MATH_2018), col="#3e3f3a")
```

```{r message=FALSE}
  boxplot(na.omit(PISA_SCI_2018), col="#3e3f3a")
```

#### Summary Statistics of the remaining Variables 

```{r echo=FALSE, message=FALSE}
  OVERALL_STAT <- summary(PISA_2018)
  OVERALL_STAT
```

### Data Cleaning

#### To-do's:
- [x] Renaming all columns to UPPERCASE and shorten its name
- [x] Omit NA values from all columns
- [x] Rebuilding new student profile

```{r message=FALSE}

# Renaming columns to UPPERCASE
CNT <- PISA_2018$country
SCH_ID <- PISA_2018$school_id
STU_ID <- PISA_2018$country
MOTHER_ED <- PISA_2018$mother_educ
FATHER_ED <- PISA_2018$father_educ
GENDER <- PISA_2018$gender
COMPUTER <- PISA_2018$computer
INTERNET <- PISA_2018$internet
MATH <- PISA_2018$math
SCIENCE <- PISA_2018$science
STU_WGT <- PISA_2018$stu_wgt
DESK <- PISA_2018$desk
NO_TV <- PISA_2018$television
NO_COMPUTER <- PISA_2018$computer_n
NO_CAR <- PISA_2018$car
NO_BOOK <- PISA_2018$book
WEALTH <- PISA_2018$wealth
ESCS <- PISA_2018$escs

# Rebuilding new student profile
# Invoking na.omit to remove missing values from the data-set
STUDENT_PROFILE_2018 = na.omit(data.frame(
  CNT,
  SCH_ID,
  STU_ID,
  MOTHER_ED,
  FATHER_ED,
  GENDER,
  COMPUTER,
  INTERNET,
  MATH,
  SCIENCE,
  STU_WGT,
  DESK,
  NO_TV,
  NO_COMPUTER,
  NO_CAR,
  NO_BOOK,
  WEALTH,
  ESCS
))

# Summarizing new student profile data-set
summary(STUDENT_PROFILE_2018)

# New student profile
SAMPLE_STUD_PROF_2018 <- sample_n(STUDENT_PROFILE_2018, 10)
reactable(SAMPLE_STUD_PROF_2018)
``` 

### Data Processing
