---
title: "Process"
---

<div class="alert alert-danger" role="alert">
  <strong> NOTE: </strong> Testing data for K-Means Clustering 🧪🧪🧪
</div>

<details open="open">
  <summary>Table of Contents</summary>
  <ol>
    <li><a href="#dependency-imports">Dependencies</a></li>
    <li><a href="#importing-pisa-2018-data">Importing Data</a></li>
    <li><a href="#exploring-the-data">Data Exploration</a></li>
    <li><a href="#license">License</a></li>
    <li><a href="#contact">Contact</a></li>
    <li><a href="#acknowledgements">Acknowledgements</a></li>
  </ol>
</details>

<hr/>

### Dependency Imports
```{r message=FALSE}
library(learningtower)
library(dplyr)
library(tidyr)
library(factoextra)
library(reactable)
```

<hr/>

### Importing PISA 2018 Data
```{r message=FALSE}

# Loading the PISA 2018 data and assigning it to variable PISA_2018
PISA_2018 <- load_student(2018)

# Sampling 10 observations from the data set for showing example
PISA_2018_SAMPLE <- sample_n(PISA_2018, 10)

# Output of the sample
reactable(PISA_2018_SAMPLE, 
  searchable = TRUE,
  striped = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  theme = reactableTheme(
    borderColor = "#dfe2e5",
    stripedColor = "#f6f8fa",
    highlightColor = "#f0f5f9",
    cellPadding = "8px 12px"))
```

#### PISA 2018 Data Format

<table class="table table-hover">
  <thead>
    <tr>
      <th>Variable Name</th>
      <th>Variable Type</th>
    </tr>
  </thead>
  <tbody>
    <tr>
      <td>year</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>country</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>school_id</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>student_id</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>mother_educ</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>father_educ</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>gender</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>computer</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>internet</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>math</td>
      <td>Numeric</td>
    </tr>
    <tr>
      <td>science</td>
      <td>Numeric</td>
    </tr>
    <tr>
      <td>read</td>
      <td>Numeric</td>
    </tr>
    <tr>
      <td>stu_wgt</td>
      <td>Numeric</td>
    </tr>
    <tr>
      <td>desk</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>room</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>dishwasher</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>television</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>computer_n</td>
      <td>Factor</td>
    </tr>
    <tr>
      <td>car</td>
      <td>Factor</td>
    </tr>
  <tbody>
</table>

<hr/>

### Exploring the Data 

```{r message=FALSE}
# Class of the data
class(PISA_2018)

# Total number of rows and columns in the data set
dim(PISA_2018)
```

<table class="table table-hover">
  <tr>
    <th>
      Class
    </th>
    <th>
      Dimension
    </th>
  </tr>
  <tr>
    <td align="center">
      data.frame
    </td>
    <td align="center">
      612004 Observations
      <br/>
      22 Variables
    </td>
  </tr>
</table>
